---
title: "HM3_r"
output: pdf_document
date: "2025-03-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Homework N2


```{r}
library(ggplot2)
library(dplyr)
library(tidyr) #For prepping data for plots
library(reshape2) #For prepping data for correlation heat maps

df_phones <- read.csv(".\\mobiles_dataset.csv")
head(df_phones)
```

## Part 1: Analytical Questions (Python & R) (I am just doing the graphs for this part explanations are in Python)

### Data Preperations

```{r}
df_preped <- df_phones %>% mutate(Launched.Price.Pakistan.PKR = Launched.Price.Pakistan.PKR * 0.0036, 
                     Launched.Price.India.INR = Launched.Price.India.INR *  0.011,
                     Launched.Price.China.CNY = Launched.Price.China.CNY * 0.14, 
                     Launched.Price.Dubai.AED = Launched.Price.Dubai.AED * 0.27)
#Fixing the POCO issue
df_preped <- df_preped %>% mutate(Company.Name = ifelse(Company.Name == 'Poco', 'POCO', Company.Name))
df_preped <- df_preped %>% filter(RAM != '8GB / 12GB') #Removing the two annoying observations
df_preped$RAM = substr(df_preped$RAM,1,nchar(df_preped$RAM)-2) #Removing the GB part from the RAM
df_preped <- df_preped %>% mutate(RAM = as.numeric(RAM)) #Converting RAM to int
head(df_preped)
```


### 1. Does battery capacity influence the launched price of a smartphone? Check this variability across all currencies. Is there any type of difference between behaviors?

```{r}
grouped_by_battery <- df_preped %>% group_by(Battery.Capacity.mAh) %>% 
  summarise(Pakistan = mean(Launched.Price.Pakistan.PKR), India = mean(Launched.Price.India.INR), 
            China = mean(Launched.Price.China.CNY), Dubai = mean(Launched.Price.Dubai.AED), 
            USA = mean(Launched.Price.USA.USD))

grouped_by_battery_longwise <- grouped_by_battery %>% 
  pivot_longer(!Battery.Capacity.mAh, names_to = "Country", values_to = "Price")
ggplot(grouped_by_battery_longwise, aes(x = Battery.Capacity.mAh, y = Price)) + 
  geom_point() +
  facet_wrap(~Country) +
  theme_bw()
```
```{r}
cormat <- round(cor(grouped_by_battery),2)
melted <- melt(cormat)
ggplot(melted, aes(x=Var1, y=Var2, fill=value)) + geom_tile() + 
  geom_text(aes(Var2, Var1, label = value), color = "black", size = 4) +
  labs(x='',y='')
```

### 2. Does RAM size impact the price of smartphones? Check this variability across all currencies. Is there any type of difference between behaviors?

```{r}
grouped_by_RAM <- df_preped %>% group_by(RAM) %>% 
  summarise(Pakistan = mean(Launched.Price.Pakistan.PKR), India = mean(Launched.Price.India.INR), 
            China = mean(Launched.Price.China.CNY), Dubai = mean(Launched.Price.Dubai.AED), 
            USA = mean(Launched.Price.USA.USD))

grouped_by_RAM_longwise <- grouped_by_RAM %>% 
  pivot_longer(!RAM, names_to = "Country", values_to = "Price")
ggplot(grouped_by_RAM_longwise, aes(x = RAM, y = Price)) + 
  geom_point() +
  facet_wrap(~Country) +
  theme_bw() +
  labs(x = 'RAM in GB')
```

```{r}
cormat <- round(cor(grouped_by_RAM),2)
melted <- melt(cormat)
ggplot(melted, aes(x=Var1, y=Var2, fill=value)) + geom_tile() + 
  geom_text(aes(Var2, Var1, label = value), color = "black", size = 4) +
  labs(x='',y='')
```
### 3. Do Apple devices have a higher price variation across different regions compared to other brands? In which country do Apple devices have the highest markup? Are there brands with more stable pricing across regions?

```{r}
pain <- df_preped %>% select(Company.Name, Launched.Price.Pakistan.PKR, Launched.Price.India.INR, 
                             Launched.Price.China.CNY, Launched.Price.USA.USD, Launched.Price.Dubai.AED)
pain <- pain %>% rename(Pak. = Launched.Price.Pakistan.PKR, Ind. = Launched.Price.India.INR,
                        Ch. = Launched.Price.China.CNY, USA = Launched.Price.USA.USD,
                        Dub. = Launched.Price.Dubai.AED)
pain_longwise <- pain %>% 
  pivot_longer(!Company.Name, names_to = "Country", values_to = "Price")

ggplot(pain_longwise, aes(x=Country, y=Price)) + geom_boxplot() + facet_wrap(~Company.Name) + theme_bw()
```



### 4. Do all smartphone brands have flagship and budget-friendly models, or do some brands only focus on premium devices? Check how many models each brand has in each segment. Determine whether a brand covers all three segments or focuses only on premium/mid-range.

```{r}
df_preped <- df_preped %>% mutate(Avg.Price = (Launched.Price.Pakistan.PKR + Launched.Price.India.INR + Launched.Price.China.CNY
                                  + Launched.Price.USA.USD + Launched.Price.Dubai.AED) / 5)  
df_preped <- df_preped %>% mutate(Device.Type = ifelse(Avg.Price < 300, 'Budget', ifelse(Avg.Price < 700, 
                                                                                         'Mid-Range', 'Premium')))
grouped_by_Comp_and_Device_Type <- df_preped %>% count(Company.Name,Device.Type,.drop=FALSE)

ggplot(grouped_by_Comp_and_Device_Type, aes(fill=Device.Type, y=n, x=Company.Name)) + 
  geom_bar(position="stack", stat="identity") +
  labs(fill = 'Type', x = 'Company', y='Count') +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 0.5)
  )
```

### 5. Which region offers the most affordable smartphone prices on average? Are there any brands that price their phones significantly lower in one region compared to others?


```{r}
df_last <- df_preped %>% summarise(Pakistan = mean(Launched.Price.Pakistan.PKR), India = mean(Launched.Price.India.INR), 
            China = mean(Launched.Price.China.CNY), Dubai = mean(Launched.Price.Dubai.AED), 
            USA = mean(Launched.Price.USA.USD))
df_last_longwise <- df_last %>% 
  pivot_longer(c(Pakistan, India, China, Dubai, USA),names_to = "Country", values_to = "Price") %>% arrange(Price)

ggplot(df_last_longwise, aes(x=reorder(Country, Price), y=Price)) + geom_bar(stat='identity') +
  labs(x = 'Country')
```


## Part 2: Visualization (Python & R)

### 1. Plot a bar chart for average price per region in USD.
```{r}
bar_df <- df_phones %>% select(c(Launched.Price.Pakistan.PKR,Launched.Price.India.INR,
                       Launched.Price.China.CNY, Launched.Price.USA.USD, Launched.Price.Dubai.AED)) %>% mutate(Launched.Price.Pakistan.PKR = Launched.Price.Pakistan.PKR * 0.0036, 
                     Launched.Price.India.INR = Launched.Price.India.INR *  0.011,
                     Launched.Price.China.CNY = Launched.Price.China.CNY * 0.14, 
                     Launched.Price.Dubai.AED = Launched.Price.Dubai.AED * 0.27) %>% summarise_all(mean)
bar_df <- data.frame(t(bar_df)) %>% `colnames<-`(c("Mean_Price"))
bar_df$Country <- rownames(bar_df)
ggplot(bar_df, aes(x=Country, y=Mean_Price)) + geom_bar(stat='identity') + 
  scale_x_discrete(labels=c("China", "Dubai", "India", "Pakistan", "USA")) +
  labs(x = 'Mean Price', y= 'Country') +
  theme_bw()
```
### 2. Create a pie chart of the market share of smartphone brands.
```{r}
#Way to many categories for a pie chart nothing can save this.
pie_df <- df_phones %>% group_by(Company.Name) %>% count()
ggplot(pie_df, aes(x = "", y = n, fill = Company.Name)) +
  geom_col() +
  geom_text(aes(label = n),
            position = position_stack(vjust = 0.2)) +
  coord_polar(theta = "y") +
  labs(fill="Brand") +
  theme_void() + 
  theme(legend.position = "bottom")
```


## Part 3: Recreate (R only)

### 1)

```{r}
ggplot(df_phones, aes(x=Company.Name, y=Launched.Price.USA.USD, fill=Company.Name)) + 
  geom_boxplot(outlier.size = 1) +
  geom_jitter(color="black", size=1, alpha=0.5) +
  labs(title='Price Distribution by Company in USA', x='Company', y='Price in USD', 
       subtitle='A boxplot showing how the price varies by compnay, with individual data points overlaid',
       fill = 'Company Name') +
  theme_minimal() +
    theme(
      plot.title = element_text(face="bold", size=12),
      plot.subtitle = element_text(face="italic", size=7),
      axis.text.x = element_text(angle = 45, vjust = 0.8, size = 6),
      axis.text.y = element_text(size = 6),
      legend.text = element_text(size=7),
      legend.title = element_text(size=9),
      legend.key.size = unit(0.37, 'cm'),
      axis.title.x = element_text(size = 8),
      axis.title.y = element_text(size = 8)
    ) 
  
```

### 2)

```{r}
ggplot(df_phones, aes(x=Battery.Capacity.mAh, y=Launched.Price.USA.USD, color=Company.Name, size=Screen.Size.inches)) +       
  geom_point() + 
  scale_size(range = c(1,3)) +
  theme_minimal() + 
  labs(title='Battery Capacity vs. Price in USA', x='Battery Capacity', y='Price', 
       subtitle='The relationship between battery capacity, price, and screen size across different smartphone brands',
       color = 'Brand') + 
  guides(size="none") + 
  theme(
    plot.title = element_text(face="bold", size=12),
    plot.subtitle = element_text(face="italic", size=7),
    axis.text.x = element_text(size = 6),
    axis.text.y = element_text(size = 6),
    legend.text = element_text(size=7),
    legend.title = element_text(size=9),
    legend.key.size = unit(0.37, 'cm'),
    axis.title.x = element_text(size = 8),
    axis.title.y = element_text(size = 8)
  )
```

### 3)

```{r}
top5_df <- df_phones %>% filter(Company.Name %in% c('Apple', 'Honor', 'Oppo', 'Samsung', 'Vivo'))
ggplot(top5_df, aes(x=Battery.Capacity.mAh, y=Launched.Price.USA.USD, shape=Company.Name, 
                    color=Screen.Size.inches, size=Screen.Size.inches)) +
  geom_point(alpha=0.6) +
  scale_size(range = c(1,3)) +
  scale_shape_manual(values = c(16, 17, 18, 15, 19)) +
  labs(title='Battery Capacity vs. Price for Top 5 Brands', x='Battery Capacity (mAh)', y='Price (USD)', 
       subtitle='Different Shapes for Each Brand, Color by Screen Size, (USA)',
       shape = 'Brand') + 
  guides(size='none', color='none') +
  theme_minimal() +
  theme(
    plot.title = element_text(face="bold", size=12),
    plot.subtitle = element_text(face="italic", size=7),
    axis.text.x = element_text(size = 6),
    axis.text.y = element_text(size = 6),
    legend.text = element_text(size=7),
    legend.title = element_text(size=9),
    legend.key.size = unit(0.37, 'cm'),
    axis.title.x = element_text(size = 8),
    axis.title.y = element_text(size = 8)
  )
```


