---
title: "HM2"
output: pdf_document
date: "2025-02-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Homework N2

## Part 3: Use the datasets provided to create graphs

```{r}
library(dplyr)


df_lung <- read.csv("C:\\Users\\Hovgr\\OneDrive\\Desktop\\DataViz\\HM2\\lung_cancer_prediction_dataset.csv")
head(df_lung, 5)

```

```{r}
df_air <- read.csv("C:\\Users\\Hovgr\\OneDrive\\Desktop\\DataViz\\HM2\\global_air_pollution_dataset.csv")
head(df_air, 5)
```

### 1. Create a Boxplot of Lung Cancer Deaths Distribution.

```{r}
library(ggplot2)
library(ggthemes)

ggplot(df_lung, aes(x = Annual_Lung_Cancer_Deaths)) + geom_boxplot() + ylim(-1,1)
```

### 2. Create a Histogram of PM2.5 AQI Values.

```{r}
ggplot(df_air, aes(x = PM2.5_AQI_Value)) + geom_histogram(bins=50)
```

### 3. Create a Density Plot of the Lung Cancer Mortality Rate.

```{r}
ggplot(df_lung, aes(x = Mortality_Rate)) + geom_density()
```

### 4. Create a Scatter Plot by generating 100 random values from both the normal and logistic distributions. The points should be brown and use theme_solarized with argument light set to false.

```{r}
x = rnorm(100)
y = rlogis(100)

ggplot() + geom_point(aes(x,y), color="brown") + labs(x = "Normal", y = "Logistic") + theme_solarized(light=FALSE)
```

## Part 4: Recreate the following graphs

### 2. Use the gpplot2 package for this graph. (Hint: Aggregate the data then merge the two datasets. Use only the necessary columns.)

```{r}
df1 <- group_by(df_air, Country) %>% summarise(PM2.5_AQI_Value = mean(PM2.5_AQI_Value))
df2 <- group_by(df_lung, Country) %>% summarise(Annual_Lung_Cancer_Deaths = sum(Annual_Lung_Cancer_Deaths))

joined_df <- inner_join(df1, df2, by="Country")

ggplot(joined_df, aes(x = PM2.5_AQI_Value, y = Annual_Lung_Cancer_Deaths, color = Country)) +
  geom_point(aes(size=Annual_Lung_Cancer_Deaths)) + 
  labs(title = "PM2.5 AQI vs. Annual Lung Cancer Deaths", x = "PM2.5 AQI Value", y = "Annual Lung Cancer Deaths") +
  geom_text(
    aes(label = ifelse(Annual_Lung_Cancer_Deaths > 500000000, Country, ''), 
    size=Annual_Lung_Cancer_Deaths), color="black", 
    fontface = "bold") +
  theme(
    plot.title = element_text(color = "darkred", face="bold", size=15),
    plot.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(color = "black", fill=NA),
    panel.background = element_rect(fill = "#f7f7f7",
                                size = 0.5, linetype = "solid"),
    panel.grid.major = element_line(size = 0.6, linetype = 'dashed',
                                color = "gray"),
    panel.grid.minor = element_line(size = 0.25, linetype = 'solid',
                                color = "#ebebeb"),
    legend.background = element_rect(fill = "#f7f7f7"),
    axis.text.x = element_text(angle = 45, vjust = 0.5, color="blue")
    )
```

### 3. Use the ggplot2 package for this graph. (Hint: use geom_jitter since y axis contains categorical data, also use the following colors: #5469f1 , #d554f1)

```{r}
filtered_df_lung <-  df_lung %>% filter(Cancer_Stage != "None", Years_of_Smoking > 0)
ggplot(filtered_df_lung, 
       aes(x = Years_of_Smoking, y=Cancer_Stage, color = Gender, shape=Gender)
  ) + 
  geom_jitter(alpha=0.7) + 
  facet_grid(~Gender) + 
  scale_color_manual(values = c("#d554f1","#5469f1")) + 
  scale_shape_manual(values = c(17,19)) +
  labs(
    title = "Lung Cancer Stage vs. Smoking Years", 
    subtitle = "Comparison by Gender",
    x = "Years of Smoking",
    y = "Cancer Stage"
    ) + 
  theme_minimal() +
  theme(
    plot.background = element_rect(fill = "#f7f7f7"),
    plot.title = element_text(face="bold", size=15),
    plot.subtitle = element_text(face="italic"),
    strip.text.x = element_text(size = 15, face="bold"),
    legend.position = "bottom",
    legend.title = element_text(size=12),
    legend.text = element_text(size=10),
    axis.title=element_text(size=14)
  )
```

### 4. Use the ggplot2 package for this graph. (Hint: use scale_fill_viridis_d(option = "plasma" to get the same colors)

```{r}
filtered_df_air <-  df_air %>% filter(Country %in% c("Brazil", "India", "Russian Federation", "Germany", "Italy", "United States of America"))
ggplot(filtered_df_air , aes(x = PM2.5_AQI_Value, fill=Country)) + 
  facet_wrap(~Country, scales = "free_y") + 
  geom_histogram(bins=50, color="black") + 
  scale_fill_viridis_d(option = "plasma") +
  labs(
    title = "PM2.5 AQI Distribution Across Countries", 
    subtitle = "Comparison of Air Pollution Levels",
    x = "PM2.5 AQI Value",
    y = "Frequency"
    ) +
  theme_minimal() +
  theme(
    plot.background = element_rect(fill = "#f7f7f7"),
    plot.title = element_text(face="bold", size=15),
    plot.subtitle = element_text(face="italic"),
    strip.text.x = element_text(face="bold"),
    legend.position = "bottom",
    legend.title = element_text(size=12),
    legend.text = element_text(size=10),
    axis.title=element_text(size=15)
  )
```
