---
title: "HM1"
output:
  pdf_document: default
  html_document: default
date: "2025-02-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Part 1: Data Cleaning and Exploration (Python & R)

## 1) Identify columns with missing values and their respective counts Drop columns where more than 50% of the data is missing (store this version as a new dataset).
```{r}
library(dplyr)


df <- read.csv("C:\\Users\\Hovgr\\OneDrive\\Desktop\\DataViz\\HM1\\crime_data.csv")
head(df, 5)


```
## 2) Identify columns with missing values and their respective counts. Drop columns where more than 50% of the data is missing (store this version as a new dataset).

```{r}
print(colSums(is.na(df))[colSums(is.na(df)) > 0])
```

```{r}
df_cleaned <- df %>% select(-all_of(names(colSums(is.na(df))[colSums(is.na(df)) > nrow(df) * 0.5])))
head(df_cleaned, 5)
```
## 3) Convert the DATE OCC column to a datetime format. Extract the year, month, and day into separate columns. Create a new column for the hour using the TIME OCC column.
```{r}
df_cleaned <- df_cleaned %>% mutate(
 DATE.OCC = as.Date(DATE.OCC, format = "%m/%d/%Y"), 
 Year = format(DATE.OCC, "%Y"),
 Month = format(DATE.OCC, "%m"),
 Day = format(DATE.OCC, "%d"),
 Hour = as.integer(TIME.OCC / 100)
)

head(df_cleaned, 5)
```

## 4) Filter the dataset for crimes that occurred in 2023. Further filter crimes with the description BURGLARY in the Crm Cd Desc column.

``` {r}
df_cleaned <- df_cleaned %>% filter(Year == 2023, Crm.Cd.Desc == "BURGLARY")
head(df_cleaned, 5)

```
## 5) Group the data by AREA NAME and calculate the total number of crimes and the average victim age. Sort the results by total crimes in descending order.
```{r}
df_cleaned %>% group_by(AREA.NAME) %>% summarise(
    count = n(), 
    mean = mean(Vict.Age, na.rm = TRUE) ) %>% arrange(desc(count)) %>% head(5) 
```

# Part 3: Further Exploration 
##1. Group the data by Month and count the number of crimes.
```{r}
df_cleaned %>% group_by(Month) %>% summarise(count = n()) 
```

## 2) Count the number of crimes where a weapon was used (Weapon Used Cd is not null).
```{r}
df %>% filter(!is.na(Weapon.Used.Cd)) %>% summarise(count = n())
```
## 3) Group the data by Premis Desc and count the number of crimes.
```{r}
df %>% group_by(Premis.Desc) %>% summarise(count = n()) %>% head(5)
```
# Part 4: Advanced Analysis
## 1. Create a new column, Severity Score, based on the following rules:
• Assign 5 points if a weapon was used (Weapon Used Cd is not null).
• Assign 3 points for crimes under BURGLARY.
• Assign 1 point for all other crimes.
• Group by AREA NAME and find the total severity score for each area.
```{r}
df <- df %>% mutate(
    Severity.Score = case_when(
      !is.na(Weapon.Used.Cd) & Crm.Cd.Desc == "BURGLARY" ~ 8,
      !is.na(Weapon.Used.Cd) ~ 5,
      Crm.Cd.Desc == "BURGLARY" ~ 3,
      TRUE ~ 1
    )
)

df %>% group_by(AREA.NAME) %>%
  summarise(total = sum(Severity.Score)) %>%
  arrange(desc(total))

```
